---
title: "Base Analysis: empirical covariance prediction"
author: "Sophie Tascedda"
date: "`r Sys.Date()`"
output:
  html_document:
    df_print: paged
params:
  num_regions: '15'
---

## Get and prepare the data

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(CovRegFC)
library(cowplot)
library(ggplot2)
library(reshape2)
library(magrittr)
library(stringr)
library(devtools)
library(rstan)
library(abind)
#library(BatchJobs)
library(reshape2)
#library(rethinking)
#library(cap)
library(dplyr)
library(RANN)
params
```

Analysis based on data from the WU-Minn HCP 1200 Subjects Data Release with four complete rfMRI runs (with 100% of collected time points).

Load subject info
```{r}
subject_info = read.csv("data/HCP_PTN820/sample_info.csv")
subject_info = subject_info[,c(
  "Subject",
  "Age",
  "Gender",
  "Acquisition",
  "PSQI_AmtSleep", # amount of sleep in hours
  "PSQI_Score" # Pittsburgh Sleep Quality Index (PSQI) Completed
  )] 
```

Define two groups: short and conventional sleepers according to the classification from Hirshkowitz et al. (2015)^[Hirshkowitz et al. (2015), National Sleep Foundation's sleep recommendations: Methodology and results summary, Sleep Health.]:

* short sleepers: average equal or less than 6 hours each night
* conventional sleepers: average between 7 and 9 hours each night

```{r}
sleep_duration = rep("undefined",nrow(subject_info))
sleep_duration[subject_info$PSQI_AmtSleep <= 6] = "short"
sleep_duration[(subject_info$PSQI_AmtSleep >= 7) & (subject_info$PSQI_AmtSleep <= 9)] = "conventional"
subject_info$sleep_duration = factor(sleep_duration)
table(subject_info$sleep_duration)
str(subject_info)
```

Load timeseries

```{r}
num_regions = as.integer(params$num_regions)
num_regions
channel_names = paste0("R",1:num_regions)
path = paste0("data/HCP_PTN820/node_timeseries/3T_HCP820_MSMAll_d",num_regions,"_ts2") 
file_names = list.files(path = path,pattern = ".txt")
file_subject_ids = strsplit(file_names,split = ".txt") %>% unlist
ts = lapply(file_subject_ids,function(Subject) {
  print(paste("reading subject:",Subject))
  full_path = paste0(path,"/",Subject,".txt")
  timeseries = read.csv(full_path,header = FALSE,sep = " ")
  timeseries$Subject = Subject
  timeseries
}) %>% do.call(rbind,.) %>% data.frame
names(ts)[1:num_regions] = channel_names
```
Merge timeseries and subject info, and separate in 2 groups. Here, "short" sleepers and "conventional" sleepers
```{r}
ts_subject_info = merge(ts,subject_info,by = "Subject")
ts_short = subset(ts_subject_info,sleep_duration=="short")
ts_short$timepoint = rep(1:(nrow(ts_short)/4),4)
ts_conventional = subset(ts_subject_info,sleep_duration=="conventional")
ts_conventional$timepoint = rep(1:(nrow(ts_conventional)/4),4)
```

```{r}
p_avg_sleep <- subset(subject_info,sleep_duration!="undefined") %>%
  ggplot( aes(x=PSQI_AmtSleep, fill=sleep_duration)) +
    geom_histogram( color="#e9ecef", alpha=0.6, position = 'identity') +
    scale_fill_manual(values=c("#69b3a2", "#404080")) +
    theme_minimal() +
    labs(fill="")
p_avg_sleep
```

### Generate our Y and X:
Y: timeseries data
X: covariate(s)

Short sleepers
```{r}
short_ids=unique(ts_short$Subject)
short_subj=subset(subject_info,sleep_duration=="short")
short_subj <- short_subj[short_subj$Subject %in% short_ids, ]
X_short_subj<-short_subj[order(short_subj$Subject),]
Y_short <- lapply(split(ts_short,ts_short$Subject), "[", TRUE, -c(17:24))
Y_short <- lapply(Y_short, "[", TRUE, -c(1))

```

Conventional sleepers
```{r}
conventional_ids=unique(ts_conventional$Subject)
conventional_subj=subset(subject_info,sleep_duration=="conventional")
conventional_subj <- conventional_subj[conventional_subj$Subject %in% conventional_ids, ]
X_conv_subj<-conventional_subj[order(conventional_subj$Subject),]
Y_conv <- lapply(split(ts_conventional,ts_conventional$Subject), "[", TRUE, -c(17:24))
Y_conv <- lapply(Y_conv, "[", TRUE, -c(1))
```

Select 1 covariate (here "PSQI_AmtSleep") and convert Y and X to matrices (required by capReg)
```{r}
X_short_subset<-subset(X_short_subj, select="PSQI_AmtSleep")
X_short_subset<-cbind(inter = 1, X_short_subset)

X_short_subset<-data.matrix(X_short_subset)
Y_short<-lapply(Y_short, function(x) data.matrix(x))
```

```{r}
X_conv_subset<-subset(X_conv_subj, select="PSQI_AmtSleep")
X_conv_subset<-cbind(inter = 1, X_conv_subset)

X_conv_subset<-data.matrix(X_conv_subset)
Y_conv<-lapply(Y_conv, function(x) data.matrix(x))
```
Center Y
```{r}
Y_c_short<- lapply(Y_short, function(x) scale(x,center = TRUE, scale = FALSE))
#Y_c_short<- lapply(Y_c_short, function(x) data.matrix(x))
Y_c_conv<- lapply(Y_conv, function(x) scale(x,center = TRUE, scale = FALSE))
#Y_c_conv<- lapply(Y_c_conv, function(x) data.matrix(x))

#Y_c_short=tibble(Y_c_short)
#Y_c_conv=tibble(Y_c_conv)
```

## BASE ANALYSIS: Use average covariance matrix for predicting covariance matrices for the missing potential outcome

On one selected cell
```{r}
source("base_analysis.R")
source("sample_covariance.R")
source("get_triangle.R")

analysis1=base_analysis(Y_c_short, Y_c_conv, channel_names, n_row=2, n_col=1, paired=TRUE)
```

On all combinations of regions
```{r}
source("base_analysis_full.R")
source("Repeated_ttest.R")

analysis2=base_analysis_full(Y_c_short, Y_c_conv, channel_names)
```

```{r}
source("get_pvalues.R")
list_pvalues=get_pvalues(analysis2$test)
list_pvalues=as.double(list_pvalues)
hist(as.numeric(list_pvalues), breaks = seq(0,1,by=0.05), main="Histogram of pvalues", col="gold")
```

```{r}
source("back_to_matrix.R")
pvalues_matrix=back_to_matrix(num_regions, channel_names, list_pvalues, triangle="LOWER")
```

```{r}
pval_mat_long = melt(pvalues_matrix, varnames = c("Rows","Cols"), na.rm = TRUE)
pval_plot = ggplot(data = pval_mat_long,aes(Cols, Rows, fill = value)) +
    geom_tile(color = "white") +
    scale_fill_stepsn(colours = c("darkblue", "red", "lightpink","cornsilk"), space = "Lab", breaks=c( 0,0.05,0.2,0.5,1), 
                        
                         name = "Pvalues") +
    theme_minimal() +
    theme(axis.text.x = element_text(angle = 90, vjust = 1, hjust = 1),
          axis.title.x=element_blank(),
          axis.title.y=element_blank(),
          legend.title=element_blank()) +
    coord_fixed() +
  scale_y_reverse()+
    ggtitle("p-values")

pval_plot
```


## KNN ANALYSIS: Use average of k nearest neighbors based on the selected covariance for the covariance matrix prediction


We have:

Y_c_short = Ys for conventional
Y_c_conv = Ys for short

X_conv, X_short: covariates

```{r}
X_conv=as.matrix(X_conv_subset[,2])
X_short=as.matrix(X_short_subset[,2])
```

```{r}
source("knn_analysis_full.r")
analysis3=knn_analysis_full(X_short, X_conv, Y_c_short, Y_c_conv, channel_names, K=5)
```

```{r}
list_pvalues2=get_pvalues(analysis3$test)
list_pvalues2=as.double(list_pvalues2)
hist(as.numeric(list_pvalues2), breaks = seq(0,1,by=0.05), main="Histogram of pvalues", col="gold")
```



```{r}
source("back_to_matrix.R")
pvalues_matrix_knn=back_to_matrix(num_regions, channel_names, list_pvalues2, triangle="LOWER")
```

```{r}
pval_mat_long2 = melt(pvalues_matrix_knn, varnames = c("Rows","Cols"), na.rm = TRUE)
pval_plot2 = ggplot(data = pval_mat_long2,aes(Cols, Rows, fill = value)) +
    geom_tile(color = "white") +
    scale_fill_stepsn(colours = c("darkblue", "red", "lightpink","cornsilk"), space = "Lab", breaks=c( 0,0.05,0.2,0.5,1), 
                        
                         name = "Pvalues") +
    theme_minimal() +
    theme(axis.text.x = element_text(angle = 90, vjust = 1, hjust = 1),
          axis.title.x=element_blank(),
          axis.title.y=element_blank(),
          legend.title=element_blank()) +
    coord_fixed() +
  scale_y_reverse()+
    ggtitle("p-values")

pval_plot2
```


## Repeat both base analysis and knn analysis using unbiased formula when averaging (divide by n-1 instead of n)










Add a new chunk by clicking the *Insert Chunk* button on the toolbar or by pressing *Ctrl+Alt+I*.

When you save the notebook, an HTML file containing the code and output will be saved alongside it (click the *Preview* button or press *Ctrl+Shift+K* to preview the HTML file).

The preview shows you a rendered HTML copy of the contents of the editor. Consequently, unlike *Knit*, *Preview* does not run any R code chunks. Instead, the output of the chunk when it was last run in the editor is displayed.
